\vspace{-2.25cm} Multimodal Abnormity-Aware Model\\for Ocular Disease Diagnosis

### 2024-08-28 Paper/Paper.tex
add OCT image sources (not complete yet)
overview.

	\pagebreak
	\tableofcontents
	\pagebreak

@@

		Ocular diseases can be diagnosed through various methods, including using optical coherence tomography (OCT) and color fundus images. Ophthalmologists usually identify ocular abnormalities to deduce the disease.  Traditionally, the diagnosis primarily depends on the professional experience and knowledge of the ophthalmologists, which may result in high misdiagnosis rate and under-utilization of medical data.  With the widespread application of artificial intelligence (AI), deep learning (DL) has made great contributions in providing support to patients in remote areas by sharing expert knowledge \autocite{Ichhpujani_Thakur_2021}.  By leveraging DL, researchers have developed auxiliary diagnosis programs to help doctors in the process. Many studies use convolutional neural network (CNN) to analyze ocular images. Some commonly used CNNs are VGG, ResNet, and Inception \autocite{daich2023artificial}.  And for segmenting images and finding abnormalities, U-net is widely used \autocite{Ronneberger_Fischer_Brox_2015}. In training CNNs, most studies use transferred learning, which consists of three steps: learning, fine-tuning, and validation.

@@

		We develop a multimodal abnormity-aware model to aid doctors in diagnosing ocular diseases. It is a multimodal model and considers both OCT and color fundus image inputs. In addition, tt simulates the decision-making process of doctors when they examine patients with ocular diseases, i.e., first identifying the abnormities, and then deducing the disease. Thus, it partly opens up the ``black box'' of DL and presents more interpretable conclusions.

@@

		The structure of the model is shown in Fig.~\ref{fig:3_parts}. The model constitutes of 2 parts: the Abnormity Models and the Diagnosis Model.

		The Abnormity Models consists of the OCT Model and the Fundus Model, which identify OCT and Fundus abnormities, respectively. Both of the Abnormity Models are CNNs with modified final FC layers. All CNNs that are used for classification tasks can be used. In this study, we consider ResNet152, ResNet50, ResNet18 and VGG16, and choose the best CNN for both Abnormity Models. The final FC layers are modified so that the size of their output vectors is equal to the number of OCT or Fundus abnormities. We also add a softmax layer after it. After an image is inputted into one of the Abnormity Models, the model returns a probability vector containing the probabilities of different abnormities.

		The Diagnosis Model consists of two stages: the Abnormity Number Stage (Stage D1) and the Disease Probability Stage (Stage D2). In Stage D1, there is a submodel for each of the diseases. Each submodel takes input from the OCT and Fundus Models, passes the input through a FC layer and a softmax layer, and determines how many abnormities that are related to the specific disease (as in Fig.~\ref{fig:criteria}) are present in an image. Then, in Stage D2, the output of all the submodels in Stage D1 passes through a FC layer and a softmax layer, and returns a disease probability vector.

		\begin{figure}[htbp]
			\centering
			\includegraphics[width=\linewidth]{Figs/Temp.png}
			\caption{OCT and Fundus abnormities}
			\vspace{0.3cm}
			\label{fig:abnormities}
		\end{figure}

		Abnormities are symptoms that can be directly observed in the images. The model is trained to identify 12 OCT abnormities (including normal) and 9 Fundus abnormities (including normal), as shown in Fig.~\ref{fig:abnormities}.

		\begin{figure}[htbp]
			\centering
			\includegraphics[width=\linewidth]{Figs/Temp.png}
			\caption{Abnormity-to-disease conversion criteria}
			\vspace{0.3cm}
			\label{fig:criteria}
		\end{figure}

		Most of the data we acquire are only labeled with abnormities. Therefore, we compose the following abnormity-to-disease conversion criteria and use these criteria to generate data for training the Diagnosis Model (see Fig.~\ref{fig:criteria}).

@@

		Fig.~\ref{fig:fusion} shows how the different parts of the ODADS are connected.

		When multiple OCT and Fundus images are inputted, they are passed through the Abnormity Models, resulting in multiple OCT and Fundus probability vectors. For each OCT abnormity, we choose the maximum probability  from the OCT probability vectors. Thus, we can get a final OCT probability vector with all the maximum probabilities. We do this for Fundus as well, and then concatenate the final OCT and Fundus vectors as an input to the Diagnosis Model.

		Stage D1 then takes the input and generates multiple abnormity number vectors. All these vectors are concatenated and inputted into Stage D2, which finally yields the probability vector of diseases.

		The results are then given to a user interface and shown to the user.

>>>>

	\pagebreak

@@

	Ocular diseases can be diagnosed through various methods, including using optical coherence tomography (OCT) and color fundus images. Ophthalmologists usually identify ocular abnormalities to deduce the disease.  Traditionally, the diagnosis primarily depends on the professional experience and knowledge of the ophthalmologists, which may result in high misdiagnosis rate and under-utilization of medical data.  With the widespread application of artificial intelligence (AI), deep learning (DL) has made great contributions in providing support to patients in remote areas by sharing expert knowledge \autocite{Ichhpujani_Thakur_2021}.  By leveraging DL, researchers have developed auxiliary diagnosis programs to help ophthalmologists in the process. Many studies use convolutional neural network (CNN) to analyze ocular images. Some commonly used CNNs are VGG, ResNet, and Inception \autocite{daich2023artificial}.  And for segmenting images and finding abnormalities, U-net is widely used \autocite{Ronneberger_Fischer_Brox_2015}. In training CNNs, most studies use transferred learning, which consists of three steps: learning, fine-tuning, and validation.

@@

	We develop the Multimodal Abnormity-Aware Model (MAAM) to aid ophthalmologists in diagnosing ocular diseases. It considers both OCT and color fundus image inputs. In addition, it simulates the decision-making process of ophthalmologists when they examine patients, i.e., first identifying the abnormities, and then deducing the disease. The model is designed in a way that partly opens up the ``black box'' of DL and presents more interpretable information for the reference of the ophthalmologists.

@@

	The structure of the MAAM is shown in Fig.~\ref{fig:3_parts}. It consists of 2 submodels: the Abnormity Models and the Diagnosis Model. We input multiple OCT and fundus images to Abnormity Models, and output the classification of abnormities as inputs to Diagnosis Model, which generates disease probabilities as final result.

	\vspace{0.5cm}

	The Abnormity Models are classifiers for abnormities, which are symptoms that can be directly observed in the images. The MAAM identifies 11 OCT abnormities and 8 Fundus abnormities, as shown in Fig.~\ref{fig:abnormities}.  Also refer to ``Appendix'' for more details.

	\begin{figure}[htbp]
		\centering
		\includegraphics[width=\linewidth]{Figs/Temp.png}
		\caption{OCT and Fundus Abnormities}
		\vspace{0.3cm}
		\label{fig:abnormities}
	\end{figure}

	The Abnormity Models also contains 2 submodels: the OCT Model and the Fundus Model, which identify OCT and fundus abnormities, respectively. Both of the submodels leverage CNNs with modified final full-connected (FC) layers. We compare the performances of 4 commonly used CNN, ResNet152, ResNet50, ResNet18 and VGG16, and choose the best one for each submodel. The final FC layers in each CNN are modified so that the size of their output vectors is equal to the number of OCT or fundus abnormities. In order to do normalization, we perform softmax operation for the FC layer. After an image is inputted into one of the submodels, the submodel returns a probability vector for all the abnormities.

	\vspace{0.5cm}

	And the Diagnosis Model consists of two stages: the Stage D1 and Stage D2.

	\vspace{0.2cm}

	In Stage D1, we determine the severity level for each disease from the probability vectors from the Abnormity Models.  We use Abnormity to Disease Deduction Criteria (shown in Fig.~\ref{fig:criteria}) as ground truth. There are multiple submodels in Stage D1 and each one corresponds to one disease. From the deduction criteria, we determine the number of abnormities for one disease and use the number to define the severity levels. For example, there are totally 5 abnormities that are present in disease ``Referable DR'', i.e., OCT abnormity ``DME'' and Fundus abnormities ``Hemorrhage'', ``Vascular Abnormity'', ``Microaneurysm'', and ``Cotton Wool Patch''.  Therefore, when taking account of healthy status, we end up with 6 severity levels for disease ``Referable DR''.

	We use fused vector as an input for each submodels in Stage D1, and have the vector pass a FC layer with softmax operation to generate severity level probability vectors.

	\begin{figure}[htbp]
		\centering
		\includegraphics[width=\linewidth]{Figs/Temp.png}
		\caption{Abnormity-to-disease Deduction Criteria}
		\vspace{0.3cm}
		\label{fig:criteria}
	\end{figure}

	\vspace{0.2cm}

	In Stage D2, we determine the final disease probability vector.  Similarly we use a fused vector from the output of submodels in Stage D1, and have the vector pass a FC layer with softmax operation.  The result presents potential diseases that can be referenced by ophthalmologist.

	\vspace{0.5cm}

	In MAAM, we leverage the fusion operation to incorporate all the results from different submodels.  As shown in Fig.~\ref{fig:fusion}, there are 3 fusion operations in MAAM.

@@

	The first fusion occurs at the output of OCT Models or Fundus Models.  In reality, multiple OCT images and Fundus images can be used for ocular disease diagnosis. Each image goes through one of the Abnormity Models and generates a probability vector. To use the probability results from all the images, a maximization fusion is implemented for all the vectors so that the fusion yeilds one OCT abnormity probability vector and one Fundus abnormity probability vector with maximum values.

	The second fusion occurs at the interface between Abnormity Model and Stage D1. A concatenation fusion is implemented for OCT abnormality probability vector and Fundus abnormality probability vector.  The fused vector works as an input to all the submodels in Stage D1.

	The third fusion occurs at the interface between Stage D1 and Stage D2.  We concatenation all the severity level vectors to feed the model in Stage D2.  The fused vector goes through the FC layer and does a softmax operation to yield the final result.
